### 2020-11-14 02_abstract/abstract_german.tex
Initialized chapters

%=============================================================================%
%                                Dissertation                                 %
%                               Manuel Schlund                                %
%                                  (c) 2020                                   %
%=============================================================================%
%                              Abstract (German)                              %
%=============================================================================%

\chapter*{Abstract (German version)}

\begin{otherlanguage}{ngerman}
  TBA.
\end{otherlanguage}