### 2020-11-14 ch04_assessment_climate_metrics/assessment_climate_metrics.tex

%=============================================================================%
%                                Dissertation                                 %
%                               Manuel Schlund                                %
%                                  (c) 2020                                   %
%=============================================================================%
%                        Assessment of climate metrics                        %
%=============================================================================%

\chapter{Assessment of Climate Metrics in CMIP6}
\label{ch:assessment_climate_metrics}

TBA.