### 2020-12-03 end_declaration_authorship/declaration_authorship.tex
Finished section on carbon cycle

%=============================================================================%
%                                Dissertation                                 %
%                               Manuel Schlund                                %
%                                  (c) 2020                                   %
%=============================================================================%
%                          Declaration of Authorship                          %
%=============================================================================%

\chapter{Declaration of Authorship}

I assure that this thesis is a result of my personal work and that no other
than the indicated aids have been used for its completion. Furthermore I assure
that all quotations and statements that have been inferred literally or in a
general manner from published or unpublished writings are marked as such.
Beyond this I assure that the work has not been used, neither completely nor in
parts, to pass any previous examination.

\vspace{20mm}

\noindent \ThePlace{}, \TheDate{}

\begin{flushright}
  $\overline{~~~~~~~~~~~~~\raisebox{0mm}[5mm]{\TheAuthor{}}~~~~~~~~~~~~~}$
\end{flushright}