### 2021-01-06 02_abstract/abstract_english.tex
Added figures for ESMValTool contributions (chapter 3)

%=============================================================================%
%                                Dissertation                                 %
%                               Manuel Schlund                                %
%                                  (c) 2020                                   %
%=============================================================================%
%                             Abstract (English)                              %
%=============================================================================%

\chapter*{Abstract (English version)}

TBA.