### 2021-01-08 ch04_assessment_climate_metrics/assessment_climate_metrics.tex
Cleaned tex files and started chapter 04

%=============================================================================%
%                                Dissertation                                 %
%                               Manuel Schlund                                %
%                                  (c) 2020                                   %
%=============================================================================%
%                        Assessment of climate metrics                        %
%=============================================================================%

\chapter{Assessment of Climate Sensitivity in \acs{CMIP}6}
\label{ch:04:assessment_climate_metrics}

In order to reduce uncertainties in multi-model climate projections, a first
important step is the assessment of the desired target variables in the
corresponding climate model ensemble. This is particularly relevant when a new
generation of \acp{ESM} has been published, which has happened in 2020 with the
\acs{CMIP}6 ensemble.