### 2021-01-11 end_acknowledgments/acknowledgments.tex
Cleaned tex files and added text to chapter 4

%=============================================================================%
%                                Dissertation                                 %
%                               Manuel Schlund                                %
%                                  (c) 2020                                   %
%=============================================================================%
%                               Acknowledgments                               %
%=============================================================================%

\chapter{Acknowledgments}

I assure that this thesis is a result of my personal work and that no other
than the indicated aids have been used for its completion. Furthermore I assure
that all quotations and statements that have been inferred literally or in a
general manner from published or unpublished writings are marked as such.
Beyond this I assure that the work has not been used, neither completely nor in
parts, to pass any previous examination.