### 2021-02-04 ch07_summary_outlook/summary_outlook.tex
Added further tables for chapter 05 and text for chapter 06

%=============================================================================%
%                                Dissertation                                 %
%                               Manuel Schlund                                %
%                                  (c) 2020                                   %
%=============================================================================%
%                             Summary and Outlook                             %
%=============================================================================%

\chapter{Summary and Outlook}
\label{ch:07:summary_outlook}

Outlook:

- Why GBRT not used for ECS?

- GBRT used for temperature anomaly based on other studies (emergent
constraints, historical T trend, weighting following Brunner et al.)