### 2021-02-15 end_acronyms/acronyms.tex
Added first figures for chapter 6

%=============================================================================%
%                                Dissertation                                 %
%                               Manuel Schlund                                %
%                                  (c) 2020                                   %
%=============================================================================%
%                                  Acronyms                                   %
%=============================================================================%

\chapter{List of Acronyms}

\begingroup
\setstretch{0.5}
\begin{acronym}
  \acro{AerChemMIP}{Aerosol Chemistry Model Intercomparison Project}
  \acro{AOGCM}{atmosphere-ocean general circulation model}
  \acro{AR}{Assessment Report}
  \acro{C4MIP}{Coupled Climate-Carbon Cycle Model Intercomparison Project}
  \acro{CESM}{Community Earth System Model}
  \acro{CFMIP}{Cloud Feedback Model Intercomparison Project}
  \acro{CMIP}{Coupled Model Intercomparison Project}
  \acro{CH4}[CH\textsubscript{4}]{methane}
  \acro{CMOR}{Climate Model Output Rewriter}
  \acro{CO}{carbon monoxide}
  \acro{CO2}[CO\textsubscript{2}]{carbon dioxide}
  \acro{CRE}{cloud radiative effect}
  \acro{CV}{cross-validation}
  \acro{DECK}{Diagnostic, Evaluation, and Characterisation of Klima}
  \acro{DJF}{December-January-February}
  \acro{DLR}{%
    Deutsches Zentrum für Luft- und Raumfahrt (German Aerospace Center)}
  \acro{ECS}{effective climate sensitivity}
  \acro{ENSO}{El Niño-Southern Oscillation}
  \acro{ESGF}{Earth System Grid Federation}
  \acro{ESM}{Earth system model}
  \acro{ESMValCore}{core functionalities of the ESMValTool}
  \acro{ESMValTool}{Earth System Model Evaluation Tool}
  \acro{GBRT}{gradient boosted regression tree}
  \acro{GHG}{greenhouse gas}
  \acro{GPP}{gross primary production}
  \acro{GSAT}{global mean near-surface air temperature}
  \acro{GtC}{gigatonnes of carbon}
  \acro{IAM}{integrated assessment model}
  \acro{IPA}{%
    Institut für Physik der Atmosphäre (Institute of Atmospheric Physics)}
  \acro{IPCC}{Intergovernmental Panel on Climate Change}
  \acro{ITCZ}{Intertropical Convergence Zone}
  \acro{JJA}{June-July-August}
  \acro{LongRunMIP}{Long Run Model Intercomparison Project}
  \acro{LTMI}{lower-tropospheric mixing index}
  \acro{MBLC}{marine boundary layer cloud}
  \acro{MDER}{multiple diagnostic ensemble regression}
  \acro{MIP}{Model Intercomparison Project}
  \acro{ML}{machine learning}
  \acro{MMM}{multi-model mean}
  \acro{NAO}{North Atlantic Oscillation}
  \acro{NCAR}{National Center for Atmospheric Research}
  \acro{netCDF}{Network Common Data Form}
  \acro{PDF}{probability density function}
  \acro{ppm}{parts per million}
  \acro{RCP}{Representative Concentration Pathway}
  \acro{RMSE}{root-mean-square error}
  \acro{RMSEP}{root-mean-square error of prediction}
  \acro{ScenarioMIP}{Scenario Model Intercomparison Project}
  \acro{SSP}{Shared Socioeconomic Pathway}
  \acro{SST}{sea surface temperature}
  \acro{TCR}{transient climate response}
  \acro{TLC}{tropical low-level cloud}
  \acro{TOA}{top of the atmosphere}
  \acro{WCRP}{World Climate Research Programme}
  \acro{WGCM}{Working Group on Coupled Modelling}
\end{acronym}
\endgroup