### 2021-02-17 end_appendix/appendix.tex

%=============================================================================%
%                                Dissertation                                 %
%                               Manuel Schlund                                %
%                                  (c) 2020                                   %
%=============================================================================%
%                                  Appendix                                   %
%=============================================================================%

\chapter{Appendix}

\begingroup

\crefalias{section}{appendix}
\numberwithin{equation}{section}
\numberwithin{figure}{section}
\numberwithin{table}{section}
\renewcommand{\thesection}{\Alph{section}}

\section{Supplementary Materials for
  \texorpdfstring{\Cref{ch:05:paper_ecs}}{Chapter \ref{ch:05:paper_ecs}}}
\label{sec:app:cmip_models_in_chapter_5}

\vspace{\fill}

\begin{table}[!h]
  \centering
  \begin{tabulary}{\columnwidth}{>{\em}L L}
    \toprule
    \ignorecolumntype{l}{Variable short name} & Description \\
    \midrule
    cl & Cloud area fraction \\
    clt & Total cloud area fraction \\
    hur & Relative humidity \\
    hus & Specific humidity \\
    pr & Precipitation \\
    rsdt & \Acf{TOA} incoming shortwave radiation \\
    rsut & \acs{TOA} outgoing shortwave radiation \\
    rsutcs & Clear-sky \acs{TOA} outgoing shortwave radiation \\
    ta & Air temperature \\
    tas & Surface air temperature \\
    tos & Sea surface temperature \\
    ts & Surface temperature \\
    va & Northward wind speed \\
    wap & Vertical velocity \\
    \bottomrule
  \end{tabulary}
  \caption{Overview of the variables used in \cref{ch:05:paper_ecs}
    (\nameref*{ch:05:paper_ecs}). More details are given in
    \cref{tab:05:overview_emergent_constraints}, which lists the variables
    used for each emergent constraint.}
  \label{tab:app:a:overview_variables}
\end{table}

\vspace{\fill}

\begin{table}[!b]
  \centering
  \begin{tabulary}{\columnwidth}{L >{\em}L L}
    \toprule
    Observational dataset & \ignorecolumntype{l}{Corresponding variables} &
    Reference \\
    \midrule
    AIRS & hur, hus & \textcite{Aumann2003} \\
    AMSRE \ac{SST} & tos & \textcite{AMSRE2011} \\
    CERES-EBAF & rsdt, rsut, rsutcs & \textcite{Loeb2018} \\
    Cloudsat/CALIPSO & cl & \textcite{Mace2009} \\
    ERA-Interim & hur, ta, va, wap & \textcite{Dee2011} \\
    GPCP & pr & \textcite{Adler2003} \\
    HadCRUT4 & tas & \textcite{Morice2012} \\
    HadISST & ts & \textcite{Rayner2003} \\
    ISCCP D-2 & clt & \textcite{Rossow1991} \\
    MLS-Aura & hur & \textcite{Beer2006} \\
    \bottomrule
  \end{tabulary}
  \caption{References for all observational datasets used in
    \cref{ch:05:paper_ecs} (\nameref*{ch:05:paper_ecs}). More details are
    given in \cref{tab:05:overview_emergent_constraints}, which lists the
    observational datasets used for each emergent constraint.}
  \label{tab:app:a:observations}
\end{table}

\begin{table}[p]
  \centering
  \begin{tabulary}{\columnwidth}{L >{$}L<{$} L}
    \toprule
    Model & \ignorecolumntype{l}{Index} & Main reference \\
    \midrule
    ACCESS1-0 & 1 & \textcite{Dix2013} \\
    ACCESS1-3 & 2 & \textcite{Dix2013} \\
    BNU-ESM & 3 & \textcite{Ji2014} \\
    CCSM4 & 4 & \textcite{Gent2011, Meehl2012} \\
    CNRM-CM5 & 5 & \textcite{Voldoire2013} \\
    CNRM-CM5-2 & 6 & \textcite{Voldoire2013} \\
    CSIRO-Mk3-6-0 & 7 & \textcite{Rotstayn2012} \\
    CanESM2 & 8 & \textcite{Arora2011} \\
    FGOALS-g2 & 9 & \textcite{Li2013} \\
    GFDL-CM3 & 10 & \textcite{Donner2011} \\
    GFDL-ESM2G & 11 & \textcite{Dunne2012} \\
    GFDL-ESM2M & 12 & \textcite{Dunne2012} \\
    GISS-E2-H & 13 & \textcite{Schmidt2006} \\
    GISS-E2-R & 14 & \textcite{Schmidt2006} \\
    HadGEM2-ES & 15 & \textcite{Collins2011} \\
    IPSL-CM5A-LR & 16 & \textcite{Dufresne2013} \\
    IPSL-CM5A-MR & 17 & \textcite{Dufresne2013} \\
    IPSL-CM5B-LR & 18 & \textcite{Dufresne2013} \\
    MIROC-ESM & 19 & \textcite{Watanabe2011} \\
    MIROC5 & 20 & \textcite{Watanabe2010} \\
    MPI-ESM-LR & 21 & \textcite{Giorgetta2013} \\
    MPI-ESM-MR & 22 & \textcite{Giorgetta2013} \\
    MPI-ESM-P & 23 & \textcite{Giorgetta2013} \\
    MRI-CGCM3 & 24 & \textcite{Yukimoto2012} \\
    NorESM1-M & 25 & \textcite{Bentsen2013, Iversen2013} \\
    bcc-csm1-1 & 26 & \textcite{Wu2014} \\
    bcc-csm1-1-m & 27 & \textcite{Wu2014} \\
    inmcm4 & 28 & \textcite{Volodin2010} \\
    \bottomrule
  \end{tabulary}
  \caption{List of \acs{CMIP}5 models used in \cref{ch:05:paper_ecs}
    (\nameref*{ch:05:paper_ecs}) alongside the main reference and the index
    used in the corresponding figures. \AdaptedFrom{Schlund2020a}.}
  \label{tab:app:a:cmip5_models}
\end{table}

\begin{table}[p]
  \centering
  \begin{tabulary}{\columnwidth}{L >{$}L<{$} L}
    \toprule
    Model & \ignorecolumntype{l}{Index} & Main reference \\
    \midrule
    ACCESS-CM2 & 29 & \textcite{Bi2013} \\
    ACCESS-ESM1-5 & 30 & \textcite{Law2017, Ziehn2017} \\
    AWI-CM-1-1-MR & 31 & \textcite{Rackow2018, Sidorenko2015} \\
    BCC-CSM2-MR & 32 & \textcite{Wu2019} \\
    BCC-ESM1 & 33 & \textcite{Wu2019} \\
    CAMS-CSM1-0 & 34 & \textcite{Rong2018} \\
    CAS-ESM2-0 & 35 & \textcite{Wang2020} \\
    CESM2 & 36 & \textcite{Danabasoglu2020} \\
    CESM2-FV2 & 37 & \textcite{Danabasoglu2020} \\
    CESM2-WACCM & 38 & \textcite{Danabasoglu2020, Gettelman2019a} \\
    CESM2-WACCM-FV2 & 39 & \textcite{Danabasoglu2020, Gettelman2019a} \\
    CMCC-CM2-SR5 & 40 & \textcite{Cherchi2019} \\
    CNRM-CM6-1 & 41 & \textcite{Voldoire2019} \\
    CNRM-CM6-1-HR & 42 & \textcite{Voldoire2019} \\
    CNRM-ESM2-1 & 43 & \textcite{Seferian2019} \\
    CanESM5 & 44 & \textcite{Swart2019} \\
    E3SM-1-0 & 45 & \textcite{Golaz2019} \\
    EC-Earth3-Veg & 46 & \textcite{Wyser2020} \\
    FGOALS-f3-L & 47 & \textcite{Guo2020, He2019, He2020} \\
    FGOALS-g3 & 48 & \textcite{Li2020} \\
    GISS-E2-1-G & 49 & \textcite{Rind2020} \\
    GISS-E2-1-H & 50 & \textcite{Rind2020} \\
    HadGEM3-GC31-LL & 51 & \textcite{Kuhlbrodt2018} \\
    HadGEM3-GC31-MM & 52 & \textcite{Williams2018} \\
    INM-CM4-8 & 53 & \textcite{Volodin2017,Volodin2017a} \\
    INM-CM5-0 & 54 & \textcite{Volodin2017,Volodin2017a} \\
    IPSL-CM6A-LR & 55 & \textcite{Boucher2020} \\
    KACE-1-0-G & 56 & \textcite{Lee2020} \\
    MCM-UA-1-0 & 57 & \textcite{Delworth2002} \\
    MIROC-ES2L & 58 & \textcite{Hajima2020} \\
    MIROC6 & 59 & \textcite{Tatebe2019} \\
    MPI-ESM-1-2-HAM & 60 & \textcite{Mauritsen2019} \\
    MPI-ESM1-2-HR & 61 & \textcite{Muller2018} \\
    MPI-ESM1-2-LR & 62 & \textcite{Mauritsen2019} \\
    MRI-ESM2-0 & 63 & \textcite{Yukimoto2019} \\
    NESM3 & 64 & \textcite{Cao2018} \\
    NorCPM1 & 65 & \textcite{Counillon2016} \\
    NorESM2-LM & 66 & \textcite{Seland2020} \\
    NorESM2-MM & 67 & \textcite{Seland2020} \\
    SAM0-UNICON & 68 & \textcite{Park2019} \\
    TaiESM1 & 69 & \textcite{Lee2020a} \\
    UKESM1-0-LL & 70 & \textcite{Sellar2019} \\
    \bottomrule
  \end{tabulary}
  \caption{As in \cref{tab:app:a:cmip5_models} but for the \acs{CMIP}6 models.
    \AdaptedFrom{Schlund2020a}.}
  \label{tab:app:a:cmip6_models}
\end{table}

\begin{table}[p]
  \centering
  \csvreader[EmergentConstraintsPart1Table]{
    ch05_paper_ecs/data/cmip5_emergent_constraints.csv}{}{
    \dataset & $\idx$ & $\ECS$ & $\BRI$ & $\COX$ & $\LIP$ & $\SHD$ & $\SHL$ &
    $\SHS$}
  \caption{All \acs{CMIP}5 models used in \cref{ch:05:paper_ecs}
    (\nameref*{ch:05:paper_ecs}) including their \acf{ECS} and \xaxis{} values
    for the emergent constraints BRI, COX, LIP, SHD, SHL and SHS. More details
    on the emergent constraints are given in
    \cref{tab:05:overview_emergent_constraints}. The specified index
    corresponds to the index used in the associated figures.
    \AdaptedFrom{Schlund2020a}.}
  \label{tab:app:a:cmip5_emergent_constraints_part1}
\end{table}

\begin{table}[p]
  \centering
  \csvreader[EmergentConstraintsPart2Table]{
    ch05_paper_ecs/data/cmip5_emergent_constraints.csv}{}{
    \dataset & $\idx$ & $\ECS$ & $\SU$ & $\TIH$ & $\TII$ & $\VOL$ & $\ZHA$}
  \caption{As in \cref{tab:app:a:cmip5_emergent_constraints_part1} but for the
    emergent constraints SU, TIH, TII, VOL and ZHA.
    \AdaptedFrom{Schlund2020a}.}
  \label{tab:app:a:cmip5_emergent_constraints_part2}
\end{table}

\begin{table}[p]
  \centering
  \csvreader[EmergentConstraintsPart1Table]{
    ch05_paper_ecs/data/cmip6_emergent_constraints.csv}{}{
    \dataset & $\idx$ & $\ECS$ & $\BRI$ & $\COX$ & $\LIP$ & $\SHD$ & $\SHL$ &
    $\SHS$}
  \caption{As in \cref{tab:app:a:cmip5_emergent_constraints_part1} but for the
    \acs{CMIP}6 models. \AdaptedFrom{Schlund2020a}.}
  \label{tab:app:a:cmip6_emergent_constraints_part1}
\end{table}

\begin{table}[p]
  \centering
  \csvreader[EmergentConstraintsPart2Table]{
    ch05_paper_ecs/data/cmip6_emergent_constraints.csv}{}{
    \dataset & $\idx$ & $\ECS$ & $\SU$ & $\TIH$ & $\TII$ & $\VOL$ & $\ZHA$}
  \caption{As in \cref{tab:app:a:cmip5_emergent_constraints_part2} but for the
  \acs{CMIP}6 models. \AdaptedFrom{Schlund2020a}.}
  \label{tab:app:a:cmip6_emergent_constraints_part2}
\end{table}

\endgroup