### 2021-02-19 ch06_paper_gpp/paper_gpp.tex
Finished section 6.1

%=============================================================================%
%                                Dissertation                                 %
%                               Manuel Schlund                                %
%                                  (c) 2020                                   %
%=============================================================================%
%                                  Paper GPP                                  %
%=============================================================================%

\chapter{Constraining Future Gross Primary Production with Machine Learning}
\label{ch:06:paper_gpp}

As argued in detail in the previous \namecref{ch:05:paper_ecs},
single-process-oriented emergent constraints do not seem to be beneficial in
constraining climate projections from the vastly complex modern-day \acp{ESM}
since they might overly simplify the problem. To overcome this issue, we
explore an alternative approach that expands the \acf{MDER} method (see
\cref{subsec:02:mder}) with a supervised \acf{ML} algorithm. In contrast to
emergent constraints, this approach considers (1) multiple process-based
predictors, (2) multi-dimensional (gridded) target variables and (3) non-linear
relationships between the predictors and the target variable. Since the new
technique relies on a large number of data points in order to train the \ac{ML}
algorithm, \ac{ECS} as a single scalar metric for each climate model is not an
appropriate target variable. As an alternative, we analyze projections of
future \acf{GPP} in this study, which constitutes the largest flux of the
terrestrial carbon uptake. As described in detail in
\cref{subsec:02:carbon_cycle_perturbations}, the global carbon cycle actively
slows down global warming by removing \ac{CO2} from the atmosphere. In
particular, the terrestrial biosphere absorbs about $30 \unit{\%}$ of the total
anthropogenic \ac{CO2} emission \autocite{Friedlingstein2020}. Thus, accurately
quantifying the future evolution of the terrestrial carbon sink is crucial for
reliable climate change projections.

Due to complex feedback processes connected to the global carbon cycle
(concentration-carbon feedback and climate-carbon feedback; see
\cref{subsec:02:carbon_cycle_perturbations}), the response of the terrestrial
carbon cycle to changes in the atmospheric \ac{CO2} concentration and climate
is a major source of uncertainty in climate projections \autocite{Bodman2013,
  Booth2012, Collins2013}. This applies for example to multi-model projections
of future \ac{GPP} of the \acs{CMIP}5 ensemble. Thus, the aim of this study is
to reduce uncertainties in multi-model projections of future \ac{GPP} in the
\acs{CMIP}5 \acs{RCP}8.5 scenario at the end of the \nth{21} century with a new
\ac{ML}-based weighting technique and appropriate observations. The new
two-step approach utilizes aspects of emergent constraints and the \ac{MDER}
technique in combination with a \ac{ML} algorithm. In the first step, we apply
an existing emergent constraint on \ac{CO2} fertilization \autocite{Wenzel2016}
to constrain the \acp{ESM}' responses to rising atmospheric \ac{CO2}
concentration using observations of the increase of the \ac{CO2} seasonal cycle
amplitude at \ac{KUM} \autocite{Keeling2005}. In a second step, we introduce a
supervised \ac{ML} algorithm based on boosting trees \autocite{Friedman2001} to
learn an empirical spatial relationship that links grid-wise future \ac{GPP} to
historical processes relevant to its simulation under present-day conditions.
In combination with observational products of the predictors, that relationship
can be used to further constrain uncertainties in the projected spatial maps of
\ac{GPP} at the end of the \nth{21} century in the \acs{RCP}8.5 scenario. We
examine both constraining the absolute \ac{GPP} and the fractional change in
\ac{GPP} as two independent approaches and target variables. Unlike univariate
linear regression used in the \ac{MDER} algorithm, the proposed
\emph{\ac{GBRT}} algorithm is able to handle multiple predictors and copes with
non-linearities in the data. \ac{GBRT} is a well-known and successful tool used
for interpolation, classification and prediction in other fields of data
science and engineering \autocite{Death2007, Elith2008}. In the context of
climate science, \ac{GBRT} was recently applied to identify the key drivers of
spatial variations of the ratio of plant transpiration to total terrestrial
evapotranspiration in \acp{ESM} \autocite{Lian2018}.

This work is already published in \textcite{Schlund2020}. For this paper, the
author of this thesis lead the writing and the analysis and implemented the
code to reproduce this study with all figures and tables using the
\ac{ESMValTool}. \Cref{sec:06:data_and_methods} of this
\namecref{ch:06:paper_gpp} provides an overview of the data and methods used in
this paper. The results are presented in \cref{sec:06:results} and
\cref{sec:06:summary_and_discussion} closes with a summary and discussion.

\section{Data and Methods}
\label{sec:06:data_and_methods}

Here we briefly describe the data and methods used in this study.
\Cref{subsec:06:step_1} reviews the method of a previously published emergent
constraint on \ac{CO2} fertilization from \textcite{Wenzel2016} that we apply
here to our study to rescale the climate model output (step 1). This includes a
discussion of the caveats that are associated with this emergent constraint
that attributes changes in the observed increase in the \ac{CO2} seasonal cycle
amplitude at \ac{KUM} entirely to the increase in atmospheric \ac{CO2}
concentration, yet other studies provide indications that part of this increase
is due to climate and land use changes \autocite{Bastos2019, Forkel2016,
  Piao2018, Zhao2016}. In \cref{subsec:06:step_2} the \ac{ML} technique used in
step 2 to constrain gridded \ac{GPP} projections is presented.
\Cref{fig:06:schematic_steps} shows an overview of our two-step approach.
\Cref{subsec:06:uncertainty_and_skill} describes how uncertainty and predictive
skill are quantified. A complete list of all \acs{CMIP}5 models used in this
study is given in \cref{tab:app:b:cmip5_models}. To apply the emergent
constraint in step 1, we need emission-driven \acs{CMIP}5 simulations. The
number of \acp{ESM} included is therefore seven, as in \textcite{Wenzel2016}.
More details on the methods and the experimental setup are given in
\cref{sec:app:si_for_paper_gpp}.

\begin{figure}[t]
  \centering
  \includegraphics[width=\LargeFigureWidth{}]{ch06_paper_gpp/figs/1.pdf}
  \caption{Schematic illustration of our two-step approach. In step 1 an
    emergent constraint by \textcite{Wenzel2016} is used to constrain the
    global mean fractional change in \acf{GPP} over the 21st century.
    Moreover, this constraint is used to rescale two different gridded target
    variables: absolute \acs{GPP} at the end of the \nth{21} century and
    fractional \acs{GPP} change over the \nth{21} century. In step 2, a
    \acl{ML} model is used to constrain these two target variables (step 2a:
    absolute \acs{GPP}; step 2b: fractional \acs{GPP} change) in two
    independent approaches. \AdaptedFrom{Schlund2020}.}
  \label{fig:06:schematic_steps}
\end{figure}

\subsection{Constraining the \texorpdfstring{\acs{CO2}}{CO2} Fertilization
  Effect (Step 1)}
\label{subsec:06:step_1}

In this \namecref{subsec:06:step_1}, we present the observational emergent
constraint on the \ac{CO2} fertilization effect by \textcite{Wenzel2016} which
we use to globally rescale the models for their individual bias in the
sensitivity of \ac{GPP} to an increase in atmospheric \ac{CO2} concentration.
Our objective here is to present a general methodology and suggest further
improvements upon this constraint that could be used in the future.
\textcite{Wenzel2016} show a correlation between the increase in the \ac{CO2}
seasonal cycle amplitude simulated by models and the projected \ac{CO2}
fertilization. When combined with the observed trends in the \ac{CO2} seasonal
cycle amplitude, the fractional \ac{GPP} change at the time of \ac{CO2}
doubling relative to pre-industrial conditions in a \ac{1BGC} is constrained.
The predictor in this study is the sensitivity of the \ac{CO2} seasonal cycle
amplitude to rising atmospheric \ac{CO2} concentrations, defined as the slope
of the linear regression between the \ac{CO2} seasonal cycle amplitude and the
annual mean atmospheric \ac{CO2} concentrations (see \cref{fig:app:b:co2} for
details). The emergent constraint is physically motivated by the hypothesis
that increasing terrestrial \ac{GPP} is a main driver for the observed changes
in the \ac{CO2} seasonal cycle amplitude \autocite{Gray2014, Keeling1996,
  Zhao2014}. This hypothesis is based on the fact that the seasonal cycle in
atmospheric \ac{CO2} concentration originates from photosynthesis and
decomposition processes: in summer, photosynthesis removes \ac{CO2} from the
atmosphere; in winter, additional \ac{CO2} is added to the atmosphere due to
the decomposition of organic matter \autocite{Keeling1995}. Thus, \ac{CO2}
fertilization leads to an increase of the \acs{CO2} seasonal cycle amplitude
due to increased \acs{CO2} removal from the atmosphere. We note that this
emergent constraint implicitly assumes that there is no temperature-driven
increase in respiration over time. Therefore, this approach most likely
overestimates the true sensitivity of \ac{GPP} to an increase in the
atmospheric \ac{CO2} concentration.

However, we note that this emergent constraint is not undisputed. In other
studies, the observed change in the \ac{CO2} seasonal cycle amplitude is
attributed to other factors. Using a dynamic global vegetation model,
\textcite{Forkel2016} find that the increase in the \ac{CO2} seasonal cycle
amplitude above $40 \unit{\degree N}$ is mainly driven by the response of
plants to global warming instead of \ac{CO2}. Similarly, \textcite{Piao2018}
show that elevated atmospheric \acs{CO2} concentration and climate change
equally contribute to the increase in the \ac{CO2} seasonal cycle in northern
high latitudes with additional smaller contributions from air-sea carbon fluxes
and land use changes. For their analysis, \textcite{Piao2018} use \ac{CO2}
records from 26 stations and nine terrestrial ecosystem models. An evaluation
of two atmospheric inversions and eleven land surface models shows contrasting
effects of \ac{CO2} fertilization (positive) and global warming (negative) on
the \ac{CO2} seasonal cycle amplitude increase in northern high latitudes with
negligible contributions from land use changes \autocite{Bastos2019}. A further
study by \textcite{Zhao2016} finds that for seven out of nine dynamic
vegetation models, the \ac{CO2} fertilization effect is the strongest driver
for the observed changes in the \ac{CO2} seasonal cycle amplitude, while the
two remaining models equally attribute this to \ac{CO2} fertilization, climate
change and land use changes. Moreover, \textcite{Winkler2019a} discuss
limitations of carbon cycle emergent constraints related to the calculation of
the predictor from observation-based data and the inter-model relationship. As
a case study, they adopt a further emergent constraint by
\textcite{Winkler2019} that uses the response of the \ac{LAI} to ambient
\ac{CO2} instead of the \ac{CO2} seasonal cycle amplitude sensitivity to
constrain the \ac{GPP} increase in the northern high latitudes. Furthermore, it
has been shown that the \acs{CMIP}5 models \autocite{Graven2013} and the TRENDY
dynamic vegetation models \autocite{Thomas2016} greatly underestimate long-term
changes in the \ac{CO2} seasonal cycle which questions their ability to
attribute changes in the \ac{CO2} seasonal cycle to different contributors.
Nevertheless, we emphasize that we here present a generic methodology and use
the results of an existing emergent constraint. Our proposed framework is
universal and could be based on other emergent constraints in the future.

In contrast to \textcite{Wenzel2016}, instead of idealized simulations we apply
the presented emergent constraint on the emission-driven \acs{RCP}8.5
simulation \autocite{Riahi2011} and examine whether it still holds for this
simulation where forcings other than \ac{CO2} also change. Moreover, we
consider the global mean \ac{GPP} change instead of the \ac{GPP} change in the
northern extratropics as in \textcite{Wenzel2016}. Due to the strong annual
mixing of \ac{CO2} in the atmosphere, the change in the \ac{CO2} seasonal cycle
amplitude at \ac{KUM} is representative for the whole globe. Thus, we apply
this emergent constraint to the global mean \ac{GPP} change. However, we note
that this approach might introduce errors since the \ac{CO2} fertilization
effect could be different for tropical and extratropical ecosystems due to
nutrient limitations and the temperature dependence of Rubisco kinetics
\autocite{CraftsBrandner2000} in the tropics. Nevertheless, we argue that we
can use the constraint if the emergent relationship still holds within our
considered framework (the \acs{RCP}8.5 scenario simulated by the \acs{CMIP}5
models).

Therefore, our target variable in step 1 is the global mean fractional
\acs{GPP} change over the \nth{21} century $f$ in the \acs{RCP}8.5 scenario,
which is calculated from the 10-year mean \ac{GPP} at the end of the \nth{21}
century in the emission‐driven fully-coupled \acs{RCP}8.5 simulations
(\range{2091}{2100}) and the 10-year mean \ac{GPP} at the end of the \nth{20}
century (\range{1991}{2000}) in the emission-driven historical simulation:
\begin{equation}
  f = \frac{\mathtxt{GPP (\range{2091}{2100})}}{
    \mathtxt{GPP (\range{1991}{2000})}} - 1.
  \label{eq:06:f}
\end{equation}
This variable is given as a percentage, where negative values correspond to a
decrease in \ac{GPP}, positive values to an increase in \ac{GPP} and a value of
$0 \unit{\%}$ to no \ac{GPP} change. The \ac{CO2} amplitude sensitivity is
calculated for the years \range{1979}{2019}, which is the full available time
range for the observed \ac{CO2} record at \ac{KUM} \autocite{Keeling2005}.
Similar \textcite{Wenzel2016}, we extract the observed \ac{CO2} amplitude from
smoothed atmospheric \ac{CO2} concentrations (using a stiff cubic spline
function plus a four-harmonics functions with linear gain) as provided by
\textcite{Keeling2005}. For the \acs{CMIP}5 models, the respective grid cell
closest to \ac{KUM} ($19.5 \unit{\degree N}$, $154.8 \unit{\degree W}$) is
extracted. Additionally, the emission-driven historical simulations are
extended with the emission-driven \acs{RCP}8.5 simulations for the years 2006
to 2019. More details including an illustration of this calculation are shown
in \cref{fig:app:b:co2}.

Since the global constraint above cannot be used directly as a predictor in the
\ac{ML} approach which operates on gridded values (see
\cref{subsec:06:step_2}), we use it as first step to globally rescale the
\ac{ESM} output. This important first step allows us to constrain the large
uncertainty in the global \ac{CO2} fertilization effect of the \acs{CMIP}5
models. In this study, two different gridded target variables are examined (see
\cref{fig:06:schematic_steps}): the monthly climatology of future absolute
\ac{GPP} at the end of the \nth{21} century (\range{2091}{2100}) in the
\acs{RCP}8.5 scenario (see step 2a) and the fractional \acs{GPP} change over
the \nth{21} century as defined in \cref{eq:06:f} (see step 2b). The global
rescaling of the gridded target variables works as follows: Let $f_m$ be the
global mean fractional \acs{GPP} change over the \nth{21} century of \ac{ESM}
$m$, $f'$ the constrained value of the global mean fractional \ac{GPP} change
and $y_{mi}$ the target variable of climate model $m$. Then, we define the
rescaled target variable $y'_{mi}$ by
\begin{equation}
  y'_{mi} = y_{mi} \cdot \frac{f'}{f_m},
  \label{eq:06:y}
\end{equation}
where the index $i$ counts over all grid cells (and months when absolute
\ac{GPP} is used as target variable in step 2a) and $m$ refers to the \ac{ESM}
($m \in \left\{ 1, 2, \ldots, 7 \right\}$). Thus, values for the individual
grid cells/months are rescaled by a factor that is determined by the ratio
between the models' global mean \ac{GPP} change and the constrained global mean
\ac{GPP} change derived from the \textcite{Wenzel2016} emergent constraint
using the observed change in the \ac{CO2} cycle amplitude. Models that simulate
a too high global mean \ac{GPP} change are nudged toward smaller values and
vice versa. The rescaled target variables are then fed into the \ac{ML} model
(step 2) presented in the following \namecref{subsec:06:step_2}. We note that
this global rescaling assumes that the relative biases in the fractional
\ac{GPP} change of the individual \acp{ESM} are constant over the whole globe.

\subsection{Gradient Boosted Regression Tree (\acs{GBRT}) Algorithm (Step 2)}
\label{subsec:06:step_2}

For step 2 we use the \ac{GBRT} algorithm to find an optimal regression
function $F$ relating the target variable $y$ to a set of $K$ predictors
$\bm{x} = \left( x^{(1)}, x^{(2)}, \ldots, x^{(K)} \right)$ (also called
\emph{features} or \emph{covariates}) such that the predicted target variable
is given by $\hat{y} = F(\bm{x})$ \autocite{Death2007, Elith2008, Friedman2001,
  Friedman2002}. Its basic elements are decision trees, which use binary splits
of the input data to create decision rules. Due to their simple nature,
decision trees are easy to interpret but cannot be used to create satisfying
predictions for very complex datasets. This issue can be overcome by a
technique called \emph{boosting} \autocite{Freund1996}. Boosting iteratively
improves the performance of \emph{weak learners} (in our case decision trees)
in the following way: At first, a single decision tree is fitted to a randomly
selected subsample of the training data $\left\{ \left( \bm{x}_i, y_i \right)
\mid i \in I \right\}$ with an arbitrary index set $I$ (using a predefined loss
function, \eg{} the mean square error). Then, further decision trees are
linearly added to the regression function $F$ (scaled by a regularization
parameter $\nu \ll 1$ called \emph{learning rate}) and the model is fitted
again in the same manner. This ensures that the prediction for poorly modeled
training points in the beginning will gradually improve during the training.
The whole procedure is repeated until the loss function does not improve
anymore, as evaluated on an independent validation data set. Relevant
hyperparameters of this algorithm (maximum number of decision trees, maximum
depth of the decision trees, learning rate and the subsample ratio) are
optimized by $5$-fold \acf{CV} using the training data. More details on the
\ac{GBRT} algorithm are given in \cref{subsec:app:b:gbrt}.

A convenient way to evaluate and interpret \ac{GBRT} models (or other
statistical models in general) is to identify the most relevant predictors.
Globally, this so-called \emph{feature importance} is closely related to the
number of appearances of a predictor in the different decision trees that set
up the \ac{GBRT} model \autocite{Friedman2001}. To evaluate the feature
importance regionally (for the different grid cells), we use a model
distillation technique called \emph{\ac{LIME}}, which approximates the
\ac{GBRT} model with a local linear regression model for each grid cell
\autocite{Ribeiro2016}. The corresponding linear coefficients $b_k$ for each
predictor $x^{(k)}$ constitute a good proxy to the model's Jacobian
\begin{equation}
  b_k \approx \frac{\partial y}{\partial x^{(k)}},
  \label{eq:06:jacobian}
\end{equation}
which is difficult to compute explicitly for this particular \ac{ML} model.
Thus, features with high (low) absolute $b_k$ have a high (low) impact on the
local prediction of the target variable $y$.

\begin{figure}[p]
  \centering
  \includegraphics[width=\LargerFigureWidth{}]{ch06_paper_gpp/figs/2.pdf}
  \caption{Schematic illustration of our \acf{ML} approach to constrain
    projected absolute \acf{GPP} in step 2a. (a) In the training phase of the
    algorithm, the model is fitted to the training data interpolating the
    empirical (non-linear) relationship between two process-oriented
    diagnostics of the past climate $\left\{ x^{(1)}, x^{(2)} \right\}$ and
    (rescaled) future \acs{GPP} (gray surface). The points show the training
    points for the supervised \acs{ML} algorithm, each of them representing a
    single grid cell/month of a single climate model (the different colors
    correspond to different climate models). (b) In the prediction phase,
    observation-based values of the diagnostic (black points) are fed into the
    trained \acs{ML} model to constrain \acs{GPP} for every grid cell/month to
    a value which best agrees with the observations. (c) For an independent
    validation of our method, we use an out-of-sample testing setup based on a
    leave-one-model-out \acl{CV} approach (see
    \cref{subsec:02:model_weighting} for details). The schematic illustration
    of step 2b differs only in the target variable used (fractional \acs{GPP}
    change instead of absolute \acs{GPP}). \AdaptedFrom{Schlund2020}.}
  \label{fig:06:mlr_concept}
\end{figure}

\begin{table}[t]
  \centering
  \begin{tabular}[t]{p{0.27\columnwidth} l p{0.11\columnwidth}
      p{0.3\columnwidth}}
    \toprule
    Predictor & \makecell{Observation‐driven \\ data} & Used time range &
    Physical connection to \acs{GPP} \\
    \midrule
    \predictor{GPP} & \makecell{FLUXNET-MTE \\ \autocite{Jung2011}} &
    \range{1991}{2000} & -- \\
    \midrule
    \predictor{LAI} & \makecell{LAI3g \\ \autocite{Zhu2013}} &
    \range{1982}{2005} & \acs{LAI} is a measure for the number of leaves in a
    grid cell. The photosynthesis rate is highly dependent on the number of
    leaves (and vegetation in general). \\
    \midrule
    \predictor{PR} & \makecell{CRU \\ \autocite{Harris2014}} &
    \range{1901}{2005} & Water is essential for the chemical processes of
    photosynthesis. \\
    \midrule
    \predictor{RSDS} & \makecell{ERA-Interim \\ \autocite{Dee2011}} &
    \range{1979}{2005} & Solar radiation is essential for the chemical
    processes of photosynthesis. \\
    \midrule
    \predictor{T} & \makecell{CRU \\ \autocite{Harris2014}} &
    \range{1901}{2005} & \acs{T} and photosynthesis rate have a common driver
    (incoming solar radiation). \\
    \bottomrule
  \end{tabular}
  \caption{Process-oriented diagnostics (also called \emph{predictors} or
    \emph{features}) used in the \acf{GBRT} model to predict the target
    variables. For step 2a (target variable: absolute \acs{GPP}), all listed
    variables are monthly climatologies of the specified time ranges in the
    historical climate. For step 2b (target variable: fractional \acs{GPP}
    change), the temporal mean over the specified time ranges is calculated
    for all variables. \AdaptedFrom{Schlund2020}.}
  \label{tab:06:predictors}
\end{table}

As already described in the previous \namecref{subsec:06:step_1}, we analyze
two different target variables in this study: the monthly climatology of
absolute \ac{GPP} at the end of the \nth{21} century (step 2a) and the
fractional \ac{GPP} change over the \nth{21} century (step 2b). A schematic
representation of step 2a is shown in \cref{fig:06:mlr_concept}. The target
variable $y$ for the \ac{ML} model is the rescaled monthly climatology of
future GPP at the end of the \nth{21} century (\range{2091}{2100}) in the
\acs{RCP}8.5 scenario given by \cref{eq:06:y}. We relate this target variable
to several process-oriented diagnostics $\bm{x}$ of the past climate
(historical simulation) listed in \cref{tab:06:predictors}. A single training
point $\left( \bm{x}_i, y_i \right)$ for the supervised \ac{ML} algorithm
corresponds to the set of the aforementioned diagnostics evaluated on a single
grid cell/month of a single climate model (see colored points in
\namecrefs{fig:06:mlr_concept} \labelcref{fig:06:mlr_concept}{\color{BrownRed}
  a} and \labelcref{fig:06:mlr_concept}{\color{BrownRed} b}). Step 2b differs
from step 2a only in the target variable used. Both steps are independent
approaches and are evaluated separately.

The main part of steps 2a and 2b each consist of two phases. In the first phase
(training of the algorithm), we build a statistical model based on the
empirical relationship between observable process-oriented variables of the
present-day climate and future \ac{GPP} in the \acs{CMIP}5 ensemble (see gray
surface in \namecrefs{fig:06:mlr_concept}
\labelcref{fig:06:mlr_concept}{\color{BrownRed} a} and
\labelcref{fig:06:mlr_concept}{\color{BrownRed} b}). This has a flavor similar
to the \ac{MDER} algorithm \commentcite{Karpechko2013, Senftleben2020,
  Wenzel2016a}{see \cref{subsec:02:mder}}. However, in contrast to the original
\ac{MDER} algorithm, our multivariate \ac{GBRT} approach is able to handle
multiple predictors with non-linear dependencies. Moreover, instead of
considering a single globally or regionally averaged value per climate model,
our approach uses spatiotemporally gridded data for every \ac{ESM} in the
ensemble. This dramatically increases the available number of training data
(even if they are not statistically independent) and uniquely enables the
statistical model to identify and exploit regional processes. Furthermore, the
large training database allows us to perform extensive out-of-sample testing,
which is crucial to ensure the \ac{ML} model's robustness.

In the second phase (predicting with the trained \ac{GBRT} model), we use
observational data products to constrain the \acs{RCP}8.5 \acs{CMIP}5
projection of \ac{GPP}. To do this, observational data for all predictors
$\bm{x}$ of the historical climate (for every grid cell/month) are fed into the
trained \ac{GBRT} model in order to get a spatially constrained projection of
\ac{GPP} (see black points in \cref{fig:06:mlr_concept}{\color{BrownRed} b}).
Similar to the \ac{MDER} algorithm, our approach assumes that there exists an
inter-model relationship between the predictors and the target variable which
also holds for the true climate. This assumption may seem weak at first glance,
but compared to the assumption of traditional weighting approaches (models that
are better in simulating the present-day climate are also better in simulating
the future climate and vice versa), the \ac{MDER} assumption is more reasonable
and allows correcting models instead of just weighing
\commentcite{Karpechko2013}{see \cref{subsec:02:mder}}. Following other
\ac{MDER} studies \autocite{Karpechko2013, Senftleben2020, Wenzel2016a}, we
also select diagnostics as predictor variables $\bm{x}$ that are related to
\ac{GPP}-relevant processes (see \cref{tab:06:predictors} for detailed physical
explanations of the links between the predictors and \ac{GPP}).

The raw \acs{CMIP}5 and observation-driven datasets are preprocessed before
entering the \ac{ML} algorithm (see \cref{subsec:app:b:data_preprocessing}). In
particular, increases greater than $300 \unit{\%}$ in the fractional \ac{GPP}
change that is used as target variable in step 2b are removed. These
unrealistically high values (making up about $5.9 \unit{\%}$ of the data) occur
in places with very small absolute \ac{GPP} values in the historical simulation
that are used in the denominator of the derivation of the fractional \ac{GPP}
change (see \cref{eq:06:f}). In addition, we randomly select $25 \unit{\%}$ of
the training data and use it as hold-out test data. This dataset is excluded
from the training and can therefore be used for an independent validation of
our statistical model.

\subsection{Assessment of Uncertainty and Predictive Skill}
\label{subsec:06:uncertainty_and_skill}

The estimation of the \acf{SPE} for the \ac{GBRT} models needs to consider
three sources of uncertainty: the uncertainty of the \ac{ML} model itself, the
uncertainty in the rescaling of the target variable (see \cref{eq:06:y}) and
the uncertainty in the input data, that is, errors in the observation-based
products. First, the \ac{SPE} of the \ac{GBRT} model itself (\ie{} uncertainty
in the position of the gray surface in \namecrefs{fig:06:mlr_concept}
\labelcref{fig:06:mlr_concept}{\color{BrownRed} a} and
\labelcref{fig:06:mlr_concept}{\color{BrownRed} b}) is estimated as the
\emph{\acf{RMSEP}}, that is, the \ac{RMSE} between the predicted and true
values of the independent hold-out test dataset \autocite{Bishop2006}. This
approach is justified when the distribution of residuals is unbiased (see
\cref{subsec:app:b:residuals}), in which case the \ac{RMSEP} equals the
standard deviation of the residuals (which we interpret as error).
\Cref{fig:app:b:residuals} shows that this is approximately correct.

The \ac{SPE} of the \ac{GBRT} model can be understood as combined error due to
the internal variability of the climate system, the different model responses
within the climate model ensemble to forcing and the incomplete description of
the climate system with the limited number of variables used as input features.
Second, the uncertainty in the rescaling is simply assessed as the uncertainty
given by the emergent constraint from step 1. Finally, the uncertainty in the
observation-based data (\ie{} uncertainty in the position of the black points
in \cref{fig:06:mlr_concept}{\color{BrownRed} b}) is assessed by error
propagation. The necessary sensitivities of the target variable to the various
predictors for each grid cell/month are given by the linear coefficients $b_k$
given by the \ac{LIME} approach introduced in \cref{subsec:06:step_2}. All
sources of uncertainty are assumed independent and are combined by adding the
squared errors, yielding a single estimate of the \ac{SPE} for each grid
cell/month. Further details on the calculation of errors are given in
\cref{subsec:app:b:uncertainty}.

The predictive power of the \ac{GBRT} algorithm and its robustness is assessed
using a \emph{leave-one-model-out \ac{CV}} approach \commentcite{Elia2002,
  Karpechko2013}{see \cref{fig:06:mlr_concept}{\color{BrownRed} c} and
  \cref{subsec:02:model_weighting}}. For this, a single climate model is removed
from the multi-model ensemble and considered as being the true climate. Our
statistical model is then fitted on all remaining climate models and a
prediction for the \enquote{true} model is created. This allows a simple
evaluation of the predictive power by calculating the \ac{RMSEP}. The whole
process is repeated for every climate model of the ensemble in order to get a
distribution of \acp{RMSEP}. Since this approach is not limited to \ac{GBRT},
it allows a simple comparison of the predictive skill of different statistical
models. We compare the \ac{GBRT} model to the simple \acs{CMIP}5 \acf{MMM} and
a \ac{LASSO} model \autocite{Tibshirani1996}. The \ac{LASSO} model is an
extended linear regression model that regularizes the least squares solution
with an L$1$ penalty term. In contrast to the ordinary least squares
regression, the \ac{LASSO} model promotes sparse solutions (\ie{} only a few
non-zero coefficients/weights), which acts as an intrinsic feature selection
and provides an effective way to combat overfitting induced by collinearity.

\section{Results}
\label{sec:06:results}

TBA.

\begin{figure}[t]
  \centering
  \begin{subfigure}[b]{0.37\columnwidth}
    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/3a.pdf}
    \caption{}
    \label{fig:06:step1:a}
  \end{subfigure}
  \begin{subfigure}[b]{0.37\columnwidth}
    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/3b.pdf}
    \caption{}
    \label{fig:06:step1:b}
  \end{subfigure}
  \begin{subfigure}[b]{0.24\columnwidth}
    \raisebox{12.9mm}{\includegraphics[width=\columnwidth]{
      ch06_paper_gpp/figs/3c.pdf}}
  \end{subfigure}
  \caption{Emergent relationship between the global mean fractional change in
    \acf{GPP} and the sensitivity of the \acs{CO2} seasonal cycle amplitude to
    rising atmospheric \acs{CO2} concentrations observed at \acf{KUM}. Colored
    points refer to \acs{CMIP}5 models, the orange line and shaded area to the
    linear regression fit and its corresponding \acl{SPE}, and the dashed black
    lines to the observational constraint. (a) Similar to
    \textcite{Wenzel2016}: The global mean fractional \acs{GPP} change after
    \acs{CO2} doubling in a \acf{1BGC} defined as 10-year mean of the
    \acs{1BGC} run centered at the time of \acs{CO2} doubling relative to the
    10-year mean of pre-industrial control conditions at the beginning of the
    \acs{1BGC} run. The sensitivity of the \acs{CO2} amplitude for the
    \acs{CMIP}5 models is calculated from the years \range{1860}{2005}, the
    corresponding observational value from the years \range{1979}{2019}. The
    constrained global mean \acs{GPP} change after \acs{CO2} doubling in the
    \acs{1BGC} run is $\pmrangeunit{30}{9}{\%}$. (b) Fractional global mean
    \acs{GPP} change over the \nth{21} century calculated from the 10-year
    mean \acs{GPP} at the end of the \nth{21} century (\range{2091}{2100}) in
    the emission-driven fully coupled \acs{RCP}8.5 simulations and the 10-year
    mean \acs{GPP} at the end of the \nth{20} century (\range{1991}{2000}) in
    the emission‐driven fully-coupled historical run (see \cref{eq:06:f}). In
    contrast to \textcite{Wenzel2016}, the sensitivity of the \acs{CO2}
    amplitude is calculated from the years \range{1979}{2019} for climate
    models and observations for better comparability (the historical
    \acs{CMIP}5 simulations are extended with the \acs{RCP}8.5 simulations for
    the years \range{2006}{2019}). The constrained global mean \acs{GPP}
    change over the \nth{21} century is $\pmrangeunit{39}{7}{\%}$.
    \AdaptedFrom{Schlund2020}.}
  \label{fig:06:step1}
\end{figure}

\begin{figure}[t]
  \centering
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \raisebox{5.5mm}{\includegraphics[width=\columnwidth]{
      ch06_paper_gpp/figs/4a.pdf}}
    \caption{}
    \label{fig:06:step2a_ml_inference:a}
  \end{subfigure}
  ~
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/4b.pdf}
    \caption{}
    \label{fig:06:step2a_ml_inference:b}
  \end{subfigure}
  \caption{(a) Box plot of the \acf{RMSEP} distributions for six different
    statistical models used to predict future absolute \acf{GPP} in step 2a
    using a leave-one-model-out \acl{CV} approach. The distribution for each
    statistical model contains seven elements (black points; one for each
    climate model used as truth) and is represented in the following way: The
    lower and upper limit of the blue boxes correspond to the $25 \unit{\%}$
    and $75 \unit{\%}$ quantiles, respectively. The central line in the box
    shows the median and the black \enquote{x} shows the mean of the
    distribution. The whiskers outside the box represent the range of the
    distribution. Compared to the \acs{CMIP}5 \acf{MMM} and its corresponding
    rescaled version (r\acs{MMM}), the prediction uncertainty measured by the
    mean \acs{RMSEP} is significantly reduced by up to $48 \unit{\%}$ and $39
    \unit{\%}$, respectively, when using other statistical models. Moreover,
    the non-linear \acf{GBRT} models can slightly reduce the mean \acs{RMSEP}
    compared to the linear \acf{LASSO} models by about $2 \unit{\%}$ for
    \acs{GBRT}-1D (using historical \acs{GPP} as single predictor) and $3
    \unit{\%}$ for the full \acs{GBRT} model (using all predictors). (b)
    Relative global feature importance for the different predictors used in
    the \acs{GBRT} model to predict future absolute \acs{GPP} (step 2a). The
    red bars correspond to positive Pearson correlation coefficients between
    all predictors and the target variable. Due to its strong positive linear
    relationship with the future \acs{GPP}, the historical \acs{GPP} is by far
    the most important predictor in the model. A local feature importance map
    using \acf{LIME} is not shown here because \acs{GPP} is the dominant
    predictor for all grid cells. \AdaptedFrom{Schlund2020}.}
  \label{fig:06:step2a_ml_inference}
\end{figure}

\begin{figure}[t]
  \centering
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/5a.pdf}
    \caption{}
    \label{fig:06:step2a_results:a}
  \end{subfigure}
  \\
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/5b.pdf}
    \caption{}
    \label{fig:06:step2a_results:b}
  \end{subfigure}
  ~
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/5c.pdf}
    \caption{}
    \label{fig:06:step2a_results:c}
  \end{subfigure}
  \caption{(a) Ratio of the rescaled \acs{CMIP}5 ensemble mean of the absolute
    \acf{GPP} at the end of the \nth{21} century (\range{2091}{2100}) in the
    \acs{RCP}8.5 scenario using \cref{eq:06:y} (r\acs{MMM}) and its unweighted
    version (\acs{MMM}). The plot shows an almost constant value over the
    whole globe with a mean of $0.92$, which corresponds to the ratio of the
    constrained global mean \acs{GPP} change over the \nth{21} century from
    step 1 ($39 \unit{\%}$) and the \acs{CMIP}5 ensemble mean global mean
    \acs{GPP} change ($43 \unit{\%}$). All values close to $0$ for the dataset
    in the denominator have been masked to avoid divisions by $0$. (b) Bias
    between r\acs{MMM} and our \acf{GBRT} prediction for the end of the
    \nth{21} century. This corresponds to step 2a of our approach. (c) Bias
    between the modeled \acs{GPP} in the \acs{CMIP}5 \acs{MMM} of the
    historical simulation and the FLUXNET-MTE observation‐based estimate of
    \acs{GPP} \autocite{Jung2011} averaged between 1991 and 2000. Over large
    swaths of the globe, the \acs{CMIP}5 ensemble overestimates \acs{GPP} (red
    color). Panels (b) and (c) show similar bias patterns (pattern correlation
    of $R^2 = 0.88$). Thus, the \acs{GBRT} prediction in step 2a is able to
    correct the historical bias of the \acs{CMIP}5 ensemble relative to the
    FLUXNET‐MTE product. \AdaptedFrom{Schlund2020}.}
  \label{fig:06:step2a_results}
\end{figure}

\begin{figure}[t]
  \centering
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/6a.pdf}
    \caption{}
    \label{fig:06:step2a_results_biases:a}
  \end{subfigure}
  ~
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/6b.pdf}
    \caption{}
    \label{fig:06:step2a_results_biases:b}
  \end{subfigure}
  \\
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/6c.pdf}
    \caption{}
    \label{fig:06:step2a_results_biases:c}
  \end{subfigure}
  ~
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/6d.pdf}
    \caption{}
    \label{fig:06:step2a_results_biases:d}
  \end{subfigure}
  \caption{Difference (a) and ratio (b) of the both biases shown in
    \cref{fig:06:step2a_results:b,fig:06:step2a_results:c}. For panel (b), all
    values close to $0$ for the dataset in the denominator have been masked to
    avoid divisions by $0$. Both panels show that our approach is only to
    first-order a bias correction, in which case both plots would only show
    constant values. (c) Comparison of the \acf{GBRT} versus \acs{GBRT}-1D
    projections of future \acf{GPP}. This panel indicates a clear difference
    between the full \acs{GBRT} model using all predictors and the \acs{GBRT}
    model using only historical \acs{GPP} as single predictor. (d) Comparison
    of the \acs{GBRT} versus the \acf{LASSO} projections of future \acs{GPP}.
    This panel shows that there is a clear difference between using the
    non-linear \acs{GBRT} model and the linear \acs{LASSO} model. The results
    of the \acs{LASSO}-1D model are not shown here because they are very
    similar to the results of the \acs{LASSO} model.
    \AdaptedFrom{Schlund2020}.}
  \label{fig:06:step2a_results_biases}
\end{figure}

\begin{figure}[p]
  \centering
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \raisebox{5mm}{\includegraphics[width=\columnwidth]{
      ch06_paper_gpp/figs/7a.pdf}}
    \caption{}
    \label{fig:06:step2b_ml_inference:a}
  \end{subfigure}
  ~
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/7b.pdf}
    \caption{}
    \label{fig:06:step2b_ml_inference:b}
  \end{subfigure}
  \\
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \raisebox{12.5mm}{\includegraphics[width=\columnwidth]{
      ch06_paper_gpp/figs/7c.pdf}}
    \caption{}
    \label{fig:06:step2b_ml_inference:c}
  \end{subfigure}
  ~
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/7d.pdf}
    \caption{}
    \label{fig:06:step2b_ml_inference:d}
  \end{subfigure}
  \caption{(a) Box plot of the \acf{RMSEP} distributions for four different
    statistical models used to predict the fractional change in \acf{GPP} over
    the \nth{21} century (step 2b) using the leave‐one‐model‐out \acl{CV}
    approach. A detailed description of the representation of these
    distributions is given in \cref{fig:06:step2a_ml_inference:a}. The
    \acf{GBRT} algorithm shows the minimal mean and median \acs{RMSEP}.
    Compared to the \acs{CMIP}5 \acf{MMM}, its corresponding rescaled version
    (r\acs{MMM}), and the linear \acf{LASSO} model, the mean \acs{RMSEP} of the
    \acs{GBRT} model is reduced by more than $16 \unit{\%}$, $9 \unit{\%}$, and
    $3 \unit{\%}$, respectively. (b) Relative global feature importance for
    the different features used in the \acs{LASSO} model to predict the
    fractional \acs{GPP} change (step 2b). The feature importance for the
    \acs{LASSO} model is given by the normalized linear coefficients of the
    model. (c) Local feature importance for the \acs{GBRT} model used to
    predict fractional \acs{GPP} change (step 2b) calculated using the
    \acf{LIME} approach \autocite{Ribeiro2016} for the three dominant features
    \acf{T}, \acf{LAI} and \acs{GPP}. The relative influence of these three
    features (ignoring all other features) is color-coded according to the
    triangle in the lower left corner. Over large parts of the globe, \acs{T}
    is the dominant feature. (d) Relative global feature importance for the
    different features used in the \acs{GBRT} model to predict the fractional
    \acs{GPP} change (step 2b). The blue bars for the two global feature
    importance plots (b) and (d) correspond to negative Pearson correlation
    coefficients between all predictors and the target variable.
    \AdaptedFrom{Schlund2020}.}
  \label{fig:06:step2b_ml_inference}
\end{figure}

\begin{figure}[t]
  \centering
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/8a.pdf}
    \caption{}
    \label{fig:06:step2b_results:a}
  \end{subfigure}
  ~
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/8b.pdf}
    \caption{}
    \label{fig:06:step2b_results:b}
  \end{subfigure}
  \\
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/8c.pdf}
    \caption{}
    \label{fig:06:step2b_results:c}
  \end{subfigure}
  ~
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/8d.pdf}
    \caption{}
    \label{fig:06:step2b_results:d}
  \end{subfigure}
  \caption{Fractional change in \acf{GPP} over the \nth{21} century (2100
    versus 2000) in the \acs{RCP}8.5 scenario for different statistical models
    (step 2b): (a) \acs{CMIP}5 \acf{MMM} of the fractional \acs{GPP} change,
    (b) rescaled \acs{CMIP}5 \acl{MMM} (r\acs{MMM}) using \cref{eq:06:y}, (c)
    linear \acf{LASSO} model and (d) \acf{GBRT} model. The geographical
    patterns from the different statistical models are very similar and show a
    higher \acs{GPP} increase in high latitudes and a lower \acs{GPP} closer
    to the equator. \AdaptedFrom{Schlund2020}.}
  \label{fig:06:step2b_results}
\end{figure}

\begin{figure}[t]
  \centering
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/9a.pdf}
    \caption{}
    \label{fig:06:step2b_results_biases:a}
  \end{subfigure}
  \\
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/9b.pdf}
    \caption{}
    \label{fig:06:step2b_results_biases:b}
  \end{subfigure}
  ~
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/9c.pdf}
    \caption{}
    \label{fig:06:step2b_results_biases:c}
  \end{subfigure}
  \caption{(a) Ratio of the rescaled \acs{CMIP}5 ensemble mean of the
    fractional change in \acf{GPP} over the \nth{21} century using
    \cref{eq:06:y} (r\acs{MMM}; see \cref{fig:06:step2b_results:b}) and its
    unweighted version (\acs{MMM}; see \cref{fig:06:step2b_results:a}). The
    plot shows an almost constant value over the whole globe with a mean of
    $0.91$, which corresponds to the ratio of the constrained global mean
    \acs{GPP} change over the \nth{21} century from step 1 ($39 \unit{\%}$)
    and the \acs{CMIP}5 ensemble mean global mean GPP change ($43 \unit{\%}$).
    All values close to $0$ for the dataset in the denominator have been
    masked to avoid divisions by $0$. Absolute (b) and relative (c)
    differences between the fractional \acs{GPP} change over the \nth{21}
    century predicted by the \acf{GBRT} model (see
    \cref{fig:06:step2b_results:d}) and r\acs{MMM} (see
    \cref{fig:06:step2b_results:b}). For panel (c), all values close to $0$ for
    the dataset in the denominator have been masked to avoid divisions by $0$.
    Both plots show a good agreement over large parts of the globe
    (corresponding to values of $0$). There are large absolute differences in
    the Sahara region and central Asia. The largest relative differences
    (except for the Sahara and Arabian Peninsula region) appear over South
    America, South Africa, the west coast of Africa, the Middle East, parts of
    Australia and western parts of the United States.
    \AdaptedFrom{Schlund2020}.}
  \label{fig:06:step2b_results_biases}
\end{figure}

\begin{figure}[t]
  \centering
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/10a.pdf}
    \caption{}
    \label{fig:06:comparison_step2a_step2b:a}
  \end{subfigure}
  ~
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/10b.pdf}
    \caption{}
    \label{fig:06:comparison_step2a_step2b:b}
  \end{subfigure}
  \\
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/10c.pdf}
    \caption{}
    \label{fig:06:comparison_step2a_step2b:c}
  \end{subfigure}
  ~
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/10d.pdf}
    \caption{}
    \label{fig:06:comparison_step2a_step2b:d}
  \end{subfigure}
  \caption{Top row: Absolute future \acf{GPP} at the end of the \nth{21}
    century (\range{2091}{2100}) calculated using step 2a (a) and step 2b (b).
    Both approaches give similar results with global averages of $169
    \unit{\GtCyr{}}$ (a) and $175 \unit{\GtCyr{}}$ (b), which are both
    consistent with the global result of step 1 of
    $\pmrangeunit{171}{12}{\GtCyr{}}$. The pattern correlation between both
    approaches is $R^2 = 0.97$. Bottom row: Absolute (c) and relative (d)
    differences between panels (a) and (b). \AdaptedFrom{Schlund2020}.}
  \label{fig:06:comparison_step2a_step2b}
\end{figure}

\section{Summary and Discussion}
\label{sec:06:summary_and_discussion}

TBA.