### 2021-02-22 ch07_summary_outlook/summary_outlook.tex
Finished chapter 06

%=============================================================================%
%                                Dissertation                                 %
%                               Manuel Schlund                                %
%                                  (c) 2020                                   %
%=============================================================================%
%                             Summary and Outlook                             %
%=============================================================================%

\chapter{Summary and Outlook}
\label{ch:07:summary_outlook}

Summary:

- Here we present methods to constrain uncertainties in multi-model projections
of the future climate.

- ESMValTool is great tool to evaluate multi-model output (thus also
projections)

- ESMValTool is expanded and also used to create analyses presented in chapters
4, 5 and 6

- We find high multi-model means and high ranges of ECS and TCR in new CMIP6
ensemble (reasons, specific numbers)

- State-of-the-art methods (emergent constraints) cannot reduce uncertainties
in CMIP6 ECS projections (single-process, more complex models, etc.)

- Add numbers for chapter 5

- Novel ML-based approach: Cannot be used on ECS (scalar values)

- Add numbers for chapter 6

Outlook:

- Why GBRT not used for ECS?

- GBRT used for temperature anomaly based on other studies (emergent
constraints, historical T trend, weighting following Brunner et al.)