### 2021-03-02 02_abstract/abstract_english.tex
Added first part of English abstract and cleaned tex files

%=============================================================================%
%                                Dissertation                                 %
%                               Manuel Schlund                                %
%                                  (c) 2020                                   %
%=============================================================================%
%                             Abstract (English)                              %
%=============================================================================%

\chapter*{Abstract (English version)}

A precise quantification of climate change is crucial in order to assess
optimal mitigation and adaptation strategies. Valuable tools in this context
are \acp{ESM}, which are state-of-the-art climate models that allow numerical
simulations of the complex physical, biological and chemical processes of the
Earth system. Of particular relevance are projections of the future climate,
which enable an assessment of future climate change. Due to the chaotic nature
of the climate system, unknowns in future emission pathways and uncertainties
in the climate models, projections of the future climate are associated with
large uncertainties. In this thesis, we analyze future climate projections from
\acp{ESM} participating in the \ac{CMIP} and reduce corresponding uncertainties
with observations.