### 2021-03-02 02_abstract/abstract_german.tex

%=============================================================================%
%                                Dissertation                                 %
%                               Manuel Schlund                                %
%                                  (c) 2020                                   %
%=============================================================================%
%                              Abstract (German)                              %
%=============================================================================%

\chapter*{Abstract (German version)}

\begin{otherlanguage}{ngerman}
  Hier ensteht die deutsche Zusammenfassung sobald die englische Version final
  ist.
\end{otherlanguage}