### 2021-03-03 02_abstract/abstract_english.tex
Added English abstract and first part of acknowledgements

%=============================================================================%
%                                Dissertation                                 %
%                               Manuel Schlund                                %
%                                  (c) 2020                                   %
%=============================================================================%
%                             Abstract (English)                              %
%=============================================================================%

\chapter*{Abstract (English version)}

A precise quantification of climate change is crucial in order to assess
optimal mitigation and adaptation strategies. Valuable tools in this context
are \acfpAbstract{ESM}, which are state-of-the-art climate models that allow
numerical simulations of the complex physical, biological and chemical
processes of the Earth system. Of particular relevance are projections of the
future climate, which enable an assessment of future climate change. Due to the
chaotic nature of the climate system, unknowns in future emission pathways and
uncertainties in the climate models, projections of the future climate are
associated with large uncertainties. In this thesis, we analyze future climate
projections from \acsp{ESM} participating in the \acfAbstract{CMIP} and reduce
corresponding uncertainties with observations.

In a first step, the climate sensitivity (\ie{} the temperature response of the
climate system to an external forcing) in the latest generation of \acsp{ESM}
from \acs{CMIP}6 is evaluated. For the \acfAbstract{ECS}, the equilibrium
temperature response that follows a doubling of the atmospheric
\acfAbstract{CO2} concentration, we find a \acfAbstract{MMM} and multi-model
range of $3.74 \unit{K}$ and $\rangeunit{1.8}{5.6}{K}$, respectively. These
values are higher than in any previous \acs{CMIP} ensemble before. Moreover, a
third of the analyzed \acs{CMIP}6 models exceeds the upper bound of the likely
\acs{ECS} range of $\rangeunit{1.5}{4.5}{K}$ assessed by the Fifth \acl{AR}
(\acs{AR}5) of the \acl{IPCC} from 2013. Similarly, the transient response of
the climate system to a doubling of \acs{CO2}, also known as \acfAbstract{TCR},
shows an inter-model range of $\rangeunit{1.3}{3.0}{K}$, whose upper bound is
again higher than the likely range assessed in \acs{AR}5 of
$\rangeunit{1.0}{2.5}{K}$. A possible reason for the increased climate
sensitivity in many \acs{CMIP}6 models is a change in the microphysical
representation of mixed-phase clouds above the Southern Ocean, which reduces
the strong negative shortwave cloud phase change feedback in this region that
is present in climate models from previous \acs{CMIP} generations.

To reduce uncertainties in \acs{ECS} projected by the \acs{CMIP}6 models, we
systematically analyze eleven published emergent constraints on \acs{ECS} that
have mostly been derived from models participating in \acs{CMIP}5, the
predecessor generation of \acs{CMIP}6. Emergent constraints are potentially
promising approaches to reduce uncertainties in climate model projections by
combining observations and output from \acsp{ESM}. The focus of this study is
on testing if these emergent constraints hold for \acsp{ESM} participating
\acs{CMIP}6. Since none of the emergent constraints considered here have been
derived on the \acs{CMIP}6 ensemble, the \acs{CMIP}6 models can be used for
cross-checking of the emergent constraints on a new model ensemble. The
application of the emergent constraints to \acs{CMIP}6 data shows a decrease in
skill and statistical significance of the emergent relationship for nearly all
constraints, with this decrease being large in many cases. Consequently, the
size of the constrained \acs{ECS} ranges ($66 \unit{\%}$ confidence intervals)
widens by $51 \unit{\%}$ on average in \acs{CMIP}6 compared to \acs{CMIP}5.
This is likely related to the increased multi-model spread of \acs{ECS} in
\acs{CMIP}6, but may in some cases also be due to spurious statistical
relationships or a too small number of models in the ensemble that the emergent
constraint was originally derived from. The corresponding best estimates given
by the emergent constraints also increase from \acs{CMIP}5 to \acs{CMIP}6 by
$12 \unit{\%}$ on average. This can be at least partly explained by the
increased number of high-\acs{ECS} models in \acs{CMIP}6 without a
corresponding change in the constraint predictors, suggesting the emergence of
new feedback processes rather than changes in strength of those previously
dominant. Our results support previous studies concluding that emergent
constraints should be based on an independently verifiable physical mechanism
and that process-based emergent constraints on \acs{ECS} should rather be
thought of as constraints for the process or feedback they are actually
targeting.

To overcome these issues of single-process-oriented emergent constraints, we
introduce an alternative approach based on \acfAbstract{ML}. Since this new
technique relies on a large number of data points in order to train the
\acs{ML} algorithm, the scalar climate sensitivity expressed as \acs{ECS} or
\acs{TCR} is not an appropriate target variable. As an alternative, we study
projections of the future \acfAbstract{GPP}, which is the largest flux of the
terrestrial carbon uptake and slows down global warming by removing \acs{CO2}
from the atmosphere. In this analysis, we combine an existing emergent
constraint on \acs{CO2} fertilization with a \acs{ML} approach to constrain the
spatial variations of multi-model \acs{GPP} projections. In the first step of
our two-step approach, we use observed changes in the \acs{CO2} seasonal cycle
at \acl{KUM} to constrain the global mean \acs{GPP} at the end of the \nth{21}
century (\range{2091}{2100}) in \acl{RCP} 8.5 simulations with \acsp{ESM}
participating in \acs{CMIP}5 to $\pmrangeunit{171}{12}{\GtCyr{}}$, compared to
the unconstrained model range $\rangeunit{156}{247}{\GtCyr{}}$. In a second
step, we use a \acs{ML} model to constrain gridded future absolute \acs{GPP}
and gridded fractional \acs{GPP} change in two independent approaches. For
this, observational data is fed into the \acs{ML} algorithm that has been
trained on \acs{CMIP}5 data to learn relationships between present-day
physically-relevant diagnostics and the target variable. In a
leave-one-model-out \acl{CV} approach, the \acs{ML} model shows superior
performance to the \acs{CMIP}5 \acs{MMM}. Our approach predicts an increased
\acs{GPP} change in northern high latitudes compared to regions closer to the
equator.