### 2021-03-03 end_acknowledgments/acknowledgments.tex

%=============================================================================%
%                                Dissertation                                 %
%                               Manuel Schlund                                %
%                                  (c) 2020                                   %
%=============================================================================%
%                               Acknowledgments                               %
%=============================================================================%

\chapter{Acknowledgments}

This work has been supported by the European Union's Horizon 2020 project
\emph{Coordinated Research in Earth Systems and Climate: Experiments,
  kNowledge, Dissemination and Outreach (CRESCENDO)} under grant agreement
number 641816, the European Union's Horizon 2020 project \emph{Climate-Carbon
  Interactions in the Coming Century (4C)} under grant agreement number 821003,
the European Research Council Synergy Grant \emph{Understanding and Modelling
  the Earth System with Machine Learning (USMILE)} under grant agreement number
855187 and the \emph{Advanced Earth System Model Evaluation for CMIP
  (EVal4CMIP)} project funded by the Helmholtz Society. I acknowledge the
\ac{WCRP}'s \ac{WGCM}, which promoted \acs{CMIP}5 and \acs{CMIP}6. I thank the
climate modeling groups for producing and making available their model output,
the \ac{ESGF} for archiving the data and providing access and the multiple
funding agencies who support \ac{CMIP} and the \ac{ESGF}. The computational
resources of the Deutsches Klima Rechenzentrum (DKRZ, Hamburg, Germany) where
the \ac{ESMValTool} is fully integrated into the \ac{ESGF} infrastructure are
kindly acknowledged.

Personal acknowledgments...