### 2021-03-30 end_acknowledgments/acknowledgments.tex

%=============================================================================%
%                            Doctoral Dissertation                            %
%                            (c) by Manuel Schlund                            %
%=============================================================================%

%=============================================================================%
% This work is licensed under a
% Creative Commons Attribution 4.0 International License.
%
% You should have received a copy of the license along with this
% work. If not, see <http://creativecommons.org/licenses/by/4.0/>.
%=============================================================================%

\chapter{Acknowledgments}

This work has been supported by the European Union's Horizon 2020 project
\emph{Coordinated Research in Earth Systems and Climate: Experiments,
  kNowledge, Dissemination and Outreach (CRESCENDO)} under grant agreement
number 641816, the European Union's Horizon 2020 project \emph{Climate-Carbon
  Interactions in the Coming Century (4C)} under grant agreement number 821003,
the European Research Council Synergy Grant \emph{Understanding and Modelling
  the Earth System with Machine Learning (USMILE)} under grant agreement number
855187, and the \emph{Advanced Earth System Model Evaluation for CMIP
  (EVal4CMIP)} project funded by the Helmholtz Society. I acknowledge the
\ac{WCRP}'s \ac{WGCM}, which promoted \acs{CMIP}5 and \acs{CMIP}6. I thank the
climate modeling groups for producing and making available their model output,
the \ac{ESGF} for archiving the data and providing access, and the multiple
funding agencies who support \ac{CMIP} and the \ac{ESGF}. The computational
resources of the Deutsches Klima Rechenzentrum (DKRZ, Hamburg, Germany) where
the \ac{ESMValTool} is fully integrated into the \ac{ESGF} infrastructure are
kindly acknowledged.

Personal acknowledgments...