### 2021-04-08 end_abbreviations/abbreviations.tex
Standardized use of indefinite articles with acronyms

%=============================================================================%
%                            Doctoral Dissertation                            %
%                            (c) by Manuel Schlund                            %
%=============================================================================%

%=============================================================================%
% This work is licensed under a
% Creative Commons Attribution 4.0 International License.
%
% You should have received a copy of the license along with this
% work. If not, see <http://creativecommons.org/licenses/by/4.0/>.
%=============================================================================%

\chapter{List of Abbreviations}

\begin{acronym}[xxxxxxxxxxxxxx] % The x's define the label column width
  \acro{1BGC}[1\%BGC]{biogeochemically coupled simulation where the %
    atmospheric \ch{CO2} concentration increases by $1 \unit{\%}$ per year}
  \acro{AerChemMIP}{Aerosol Chemistry Model Intercomparison Project}
  \acro{AI}{artificial intelligence}
  \acro{AOGCM}{atmosphere-ocean general circulation model}
  \acro{AR}{Assessment Report}
  \acro{C4MIP}{Coupled Climate-Carbon Cycle Model Intercomparison Project}
  \acro{CESM}{Community Earth System Model}
  \acro{CFMIP}{Cloud Feedback Model Intercomparison Project}
  \acro{CH4}[\ch{CH4}]{methane}
  \acro{CMIP}{Coupled Model Intercomparison Project}
  \acro{CMOR}{Climate Model Output Rewriter}
  \acro{CO}{carbon monoxide}
  \acro{CO2}[\ch{CO2}]{carbon dioxide}
  \acro{CRE}{cloud radiative effect}
  \acro{CV}{cross-validation}
  \acro{DECK}{Diagnostic, Evaluation, and Characterization of Klima}
  \acro{DJF}{December-January-February}
  \acro{ECS}{effective climate sensitivity}
  \acro{ENSO}{El Niño--Southern Oscillation}
  \acro{ESGF}{Earth System Grid Federation}
  \acro{ESM}{Earth system model}
  \acro{ESMValCore}{core functionalities of the ESMValTool}
  \acro{ESMValTool}{Earth System Model Evaluation Tool}
  \acro{GBRT}{gradient boosted regression tree}
  \acro{GHG}{greenhouse gas}
  \acro{GPP}{gross primary production}
  \acro{GSAT}{global mean near-surface air temperature}
  \acro{GtC}{gigatonnes of carbon}
  \acro{HadGEM}{Hadley Centre Global Environmental Model}
  \acro{IAM}{integrated assessment model}
  \acro{IPCC}{Intergovernmental Panel on Climate Change}
  \acro{ITCZ}{Intertropical Convergence Zone}
  \acro{JJA}{June-July-August}
  \acro{KUM}{Cape Kumukahi, Hawaii}
  \acro{LAI}{leaf area index}
  \acro{LASSO}{least absolute shrinkage and selection operator}
  \acro{LIME}{local interpretable model-agnostic explanations}
  \acro{LongRunMIP}{Long Run Model Intercomparison Project}
  \acro{LTMI}{lower-tropospheric mixing index}
  \acro{MBLC}{marine boundary layer cloud}
  \acro{MDER}{multiple diagnostic ensemble regression}
  \acro{MIP}{Model Intercomparison Project}
  \acro{ML}{machine learning}
  \acro{MMM}{multi-model mean}
  \acro{N2O}[\ch{N2O}]{nitrous oxide}
  \acro{NAO}{North Atlantic Oscillation}
  \acro{NBP}{net biome production}
  \acro{NCAR}{National Center for Atmospheric Research}
  \acro{NEP}{net ecosystem production}
  \acro{NetCDF}{Network Common Data Form}
  \acro{NPP}{net primary production}
  \acro{PDF}{probability density function}
  \acro{ppm}{parts per million}
  \acro{PR}{precipitation}
  \acro{RCP}{Representative Concentration Pathway}
  \acro{RMSE}{root mean square error}
  \acro{RMSEP}{root mean square error of prediction}
  \acro{RSDS}{incoming solar radiation at surface}
  \acro{RuBisCO}{ribulose-1,5-bisphosphate carboxylase/oxygenase}
  \acro{ScenarioMIP}{Scenario Model Intercomparison Project}
  \acro{SPE}{standard prediction error}
  \acro{SSP}{Shared Socioeconomic Pathway}
  \acro{SST}{sea surface temperature}
  \acro{T}{near-surface air temperature}
  \acro{TCR}{transient climate response}
  \acro{TLC}{tropical low-level cloud}
  \acro{TOA}{top of the atmosphere}
  \acro{WCRP}{World Climate Research Programme}
  \acro{WGCM}{Working Group on Coupled Modelling}
\end{acronym}

% Special indefinite articles
\acrodefindefinite{ECS}{an}{an}
\acrodefindefinite{ML}{an}{a}
\acrodefindefinite{RMSE}{an}{a}
\acrodefindefinite{RMSEP}{an}{a}

% Special plurals
\acrodefplural{GHG}{greenhouse gases}