### 2021-04-13 ch01_introduction/introduction.tex
Small optimizations in introduction

%=============================================================================%
%                            Doctoral Dissertation                            %
%                            (c) by Manuel Schlund                            %
%=============================================================================%

%=============================================================================%
% This work is licensed under a
% Creative Commons Attribution 4.0 International License.
%
% You should have received a copy of the license along with this
% work. If not, see <http://creativecommons.org/licenses/by/4.0/>.
%=============================================================================%

\chapter{Introduction}
\label{ch:01:introduction}

\section{Motivation}
\label{sec:01:motivation}

Climate change is one of the greatest challenges for humankind today. The
warming of the climate system is \enquote{unequivocal}, and \enquote{many of
  the observed changes are unprecedented over decades to millennia}
\autocite{IPCC2014}. The changing climate increases the \enquote{likelihood of
  severe, pervasive and irreversible impacts for people, species and
  ecosystems} with \enquote{mostly negative impacts for biodiversity, ecosystem
services and economic development}, and amplifies \enquote{risks for
livelihoods and for food and human security} \autocite{IPCC2014}. Potential
drivers for climate change are all natural and anthropogenic substances and
processes that may alter the Earth's energy budget. Human influence on the
climate system is clear: Ever-increasing emissions of \acp{GHG} since the end
of the pre-industrial era largely driven by economic and population growth
led to atmospheric concentrations of \ac{CO2}, \ac{CH4}, and \ac{N2O} that
are \enquote{unprecedented in at least the last 800,000 years}
\autocite{IPCC2014}. The effects of these \ac{GHG} emissions and other
anthropogenic drivers have been \enquote{detected throughout the climate
system and are extremely likely to have been the dominant cause of the
observed warming since the mid-\nth{20} century} \autocite{IPCC2014}. The
greenhouse effect is based on the optical properties of the \acp{GHG}: While
the atmosphere is mostly transparent for the incoming solar (shortwave)
radiation, the outgoing infrared (longwave) radiation that is reflected from
the planet's surface is partly absorbed by the \ac{GHG} molecules through
excitation of their corresponding vibrational modes and re-emitted
isotropically. This process traps energy near the surface and leads to a
warming of the Earth's surface and the lower atmosphere. The individual
impact of the different drivers of climate change is measured with the
so-called \emph{radiative forcing}, which quantifies the change in energy
fluxes caused by changes in these drivers relative to pre-industrial
conditions \autocite{IPCC2013}. The total radiative forcing is positive
(which corresponds to a warming) and its largest contribution is caused by
the increase in the atmospheric concentration of \ac{CO2} due to
anthropogenic fossil fuel emissions since the mid-\nth{18} century
\autocite{IPCC2014}. Apart from their physical warming effect on the climate,
carbon-based \acp{GHG} like \ac{CO2} also directly influence the global
carbon cycle, an important biogeochemical cycle of the Earth. A crucial flux
of the carbon cycle is the \emph{\ac{GPP}}, which describes the carbon uptake
of the terrestrial biosphere due to photosynthesis. Since this land carbon
uptake absorbs about $30 \unit{\%}$ of the anthropogenic \ac{CO2} emissions
in today's climate \autocite{Friedlingstein2020}, this process substantially
slows down global warming. Other important anthropogenic drivers of climate
change are the emission of aerosols, and land-use/land-cover changes. Apart
from anthropogenic drivers, there are also natural processes that impact the
climate system like changes in solar activity or volcanic eruptions. However,
there is clear evidence that these natural drivers alone cannot explain the
observed climate change \autocite{Haustein2017}.

To successfully mitigate the massive impacts of climate change, a first
important step is the understanding of climate change and its accurate
quantification. Extremely valuable tools in this context are climate models,
which allow us to simulate the behavior of the climate system under arbitrary
conditions without having to perform (ethnically questionable) experiments in
the real world. All around the world climate research institutes develop a
variety of different climate models. Many of them participate in the \ac{CMIP},
which was initiated in 1995 by the \ac{WGCM} of the \ac{WCRP} to
\enquote{better understand past, present and future climate changes arising
  from natural, unforced variability or in response to changes in radiative
  forcing in a multi-model context} \autocite{WCRP2020}. The \ac{CMIP} models
provide crucial input for the international climate assessments given by the
\acp{AR} of the \ac{IPCC}. For example, the latest generation of climate models
from the most recent (sixth) phase of \ac{CMIP} \commentcite{Eyring2016}{known
  as \acs{CMIP}6} support the assessment of the upcoming \ac{IPCC} Sixth
\acl{AR} (\acs{AR}6), and their predecessor models from \acs{CMIP}5
\autocite{Taylor2012} have been assessed as part of the Fifth \acl{AR}
(\acs{AR}5) in 2013 \autocite{Flato2013}. Modern-day climate models, which
allow the simulation of biological and chemical processes in addition to the
dynamics of the physical components of the Earth system, are also known as
\emph{\acp{ESM}} and provide the most sophisticated simulations of the Earth's
climate. In this thesis, the terms \enquote{climate model} and
\enquote{\acl{ESM}} are used interchangeably since most modern models
participating in \ac{CMIP} are \acp{ESM} or at least have \ac{ESM} versions.

Simulations that extrapolate the state of the climate system into the future
are called \emph{climate model projections}. These include idealized
simulations with only a prescribed change in the atmospheric \ac{CO2}
concentration (\eg{} an instantaneous doubling of \ac{CO2} or \iac{CO2}
increase of $1 \unit{\%}$ per year) as well as more realistic projections that
consider different future scenarios (\eg{} a fossil fuel--based future or a
scenario that is based on sustainable development). Multi-model projections
from \ac{CMIP} show a large inter-model range in many variables that are
relevant for climate change \autocite{Collins2013, Flato2013}. A crucial and
policy-relevant example of this is climate sensitivity, which refers to the
change in the \ac{GSAT} that results from a change in the radiative forcing.
Common metrics for this are the \emph{\ac{ECS}}, which describes the
equilibrium response of the climate system after a doubling of the atmospheric
\ac{CO2} concentration, and the \emph{\ac{TCR}}, which describes the transient
response of the system to \iac{CO2} doubling. In \acs{AR}5, the assessed range
of both quantities is large with $\rangeunit{1.5}{4.5}{K}$ for \ac{ECS} and
$\rangeunit{1.0}{2.5}{K}$ for \ac{TCR} \autocite{Stocker2013}. The
corresponding inter-model ranges from the \acs{CMIP}5 models show similar
results \autocite{Flato2013}. For this reason, a careful statistical evaluation
and further refinement of the output of multi-model climate projections are
necessary to reduce associated uncertainties. A state-of-the-art technique for
this is the \emph{emergent constraints} method, which uses a physically based
inter-model relationship between an observable quantity of the Earth system and
a target variable to reduce uncertainties in the target variable with
observations \autocite{Allen2002}. An alternative approach is the weighting of
climate models based on their performance (\ie{} the distance of one model to
observational products) and interdependence (\ie{} the distance of one model to
other climate models) \autocite{Knutti2017a}. These techniques form the
baseline for the new analyses and results presented in this thesis, which
partly utilize methods from a new emerging research field in climate sciences:
\ac{AI} and \ac{ML}.

\section{Key Science Questions}
\label{sec:01:key_science_questions}

This thesis aims to reduce uncertainties in multi-model climate projections
with observations by addressing the following three
\lcnamecrefs{enum:01:question_1}:

\begingroup
\crefalias{enumi}{question}
\begin{enumerate}
  \item \label{enum:01:question_1} \KeyScienceQuestionOne{}
  \item \label{enum:01:question_2} \KeyScienceQuestionTwo{}
  \item \label{enum:01:question_3} \KeyScienceQuestionThree{}
\end{enumerate}
\endgroup

\section{Structure of the Thesis}
\label{sec:01:structure}

Parts of this thesis are already published in multiple peer-reviewed
publications (two first-author studies and six co-author studies). A complete
list of these publications is given on \cpageref{ch:bib:own}. Wherever material
from these studies is presented in this thesis, the pronoun \enquote{we} is
used to increase readability by avoiding the passive voice and to acknowledge
all involved contributors. However, unless stated otherwise, all content from
these publications (text, figures, and tables) shown in this thesis originates
from the author of this thesis. A detailed list of contributions to these
studies is given in the corresponding \lcnamecrefs{ch:03:esmvaltool}.

This thesis is structured as follows: \Cref{ch:02:scientific_background}
introduces the scientific background. This includes relevant literature that is
used as a baseline for this thesis. \Cref{ch:03:esmvaltool} gives an overview
of the contributions made to the \ac{ESMValTool}, an open-source software
package for the analysis of \acp{ESM}. These contributions helped to improve
the routine evaluation of \acp{ESM} which is useful for the entire scientific
community and led to co-authorship in four peer-reviewed studies
\autocite{Eyring2020, Lauer2020, Righi2020, Weigel2020}.
\Cref{ch:04:papers_ecs_tcr_assessment} covers the assessment of climate
sensitivity metrics like the \ac{ECS} or \ac{TCR} in the latest generation of
\acp{ESM} from \acs{CMIP}6. This work is already published in two scientific
publications \autocite{Bock2020, Meehl2020}. Since the \ac{ECS} and \ac{TCR}
are considerably higher in this new climate model generation,
\cref{ch:05:paper_ecs} describes the assessment of emergent constraints on the
\ac{ECS} for these \acs{CMIP}6 models and compares these to results derived
from \acs{CMIP}5 models. The content of this \lcnamecref{ch:05:paper_ecs} is
published in \emph{Earth System Dynamics} \autocite{Schlund2020a}.
\Cref{ch:06:paper_gpp} focuses on a new method to reduce uncertainties in
multi-dimensional (gridded) multi-model projections of the future climate with
observations based on \ac{ML}. As an example, the method is applied to \ac{GPP}
at the end of the \nth{21} century, which is already published in the
\emph{Journal of Geophysical Research: Biogeosciences} \autocite{Schlund2020}.
Finally, \cref{ch:07:conclusion} provides a summary of the results of this
thesis and gives an outlook of possible future work.