### 2021-08-03 end_acknowledgments/acknowledgments.tex
Prepared document for publication

%=============================================================================%
%                            Doctoral Dissertation                            %
%                            (c) by Manuel Schlund                            %
%=============================================================================%

%=============================================================================%
% This work is licensed under a
% Creative Commons Attribution 4.0 International License.
%
% You should have received a copy of the license along with this
% work. If not, see <http://creativecommons.org/licenses/by/4.0/>.
%=============================================================================%

\chapter{Acknowledgments}

This work has been supported by the European Union's Horizon 2020 project
\emph{Coordinated Research in Earth Systems and Climate: Experiments,
  kNowledge, Dissemination and Outreach (CRESCENDO)} under Grant Agreement
Number 641816, the European Union's Horizon 2020 project \emph{Climate-Carbon
  Interactions in the Coming Century (4C)} under Grant Agreement Number 821003,
the European Research Council Synergy Grant \emph{Understanding and Modelling
  the Earth System with Machine Learning (USMILE)} under Grant Agreement Number
855187, and the \emph{Advanced Earth System Model Evaluation for CMIP
  (EVal4CMIP)} project funded by the Helmholtz Society. I acknowledge the
\ac{WCRP}'s \ac{WGCM}, which promoted \acs{CMIP}5 and \acs{CMIP}6. I thank the
climate modeling groups for producing and making available their model output,
the \ac{ESGF} for archiving the data and providing access, and the multiple
funding agencies who support \ac{CMIP} and the \ac{ESGF}. The computational
resources of the German Climate Computing Center (DKRZ) where the
\ac{ESMValTool} is fully integrated into the \ac{ESGF} infrastructure are
kindly acknowledged.

\vspace{\baselineskip}

First of all, I would like to express my sincere gratitude to my supervisor and
primary examiner, \ThePrimaryExaminer{}. Without her continuous support,
advice, enthusiasm, expertise, and patience, this \PhD{} project would not have
been possible. I further thank her for providing me the opportunity to visit
many international conferences, which allowed me to meet amazing people and
collaborate with colleagues from all around the world.

I am grateful to my secondary examiner, \TheSecondaryExaminer{}, for his great
advice and feedback on my work.

My sincere thanks go to all my co-authors, namely \Prof{} \Dr{} Gustau
Camps-Valls, \ThePrimaryExaminer{}, \Prof{} \Dr{} Pierre Friedlingstein,
\TheSecondaryExaminer{}, \Dr{} Axel Lauer, \Prof{} \Dr{} Markus Reichstein, and
\Prof{} \Dr{} Steven Sherwood, for sharing their expertise, their advice, and
helpful comments on my studies.

I would like to thank \Prof{} \Dr{} Gustau Camps-Valls and his Image and Signal
Processing (ISP) group for hosting me during my research stay at the University
of Valencia. I really enjoyed the time there, where I had the opportunity to
meet great people and expand my knowledge about \acl{ML}.

Many thanks to \Dr{} Axel Lauer for being an amazing second author and for the
endless patience he showed whenever I had questions for him. I am also very
grateful to \Dr{} Birgit Hassler for her advice in all circumstances and for
providing the urgently-needed chocolate and cookies during long hours in the
office. A warm thank you to all my colleagues at DLR-EVA and IUP for their
professional advice, helpful discussions, and simply the great time we spent
together inside and outside of the office.

A special thank you to the \ac{ESMValTool} core development team for providing
me useful insights into modern software development and for the great workshops
and beers we shared.

I am deeply grateful to my friends for all the amazing time we spent together
in the many years of our friendship, which allowed me to stay focused on my
work.

A very special thanks to the (extended) Langer family for their endless support
and advice in all situations of life.

I am forever grateful to my parents and my sister for their everlasting support
and their unconditional love, without which this work would not have been
possible.

Finally, I would like to thank Hannah with all my heart, for everything.