### 2021-08-04 00_header/metadata.tex
Use PDF/A-2b format for publication

%=============================================================================%
%                            Doctoral Dissertation                            %
%                            (c) by Manuel Schlund                            %
%=============================================================================%

%=============================================================================%
% This work is licensed under a
% Creative Commons Attribution 4.0 International License.
%
% You should have received a copy of the license along with this
% work. If not, see <http://creativecommons.org/licenses/by/4.0/>.
%=============================================================================%

% Titles
\newcommand*{\Prof}{Prof.\@}
\newcommand*{\Dr}{Dr.\@}
\newcommand*{\PhD}{Ph.D.\@}

% Dissertation information
\newcommand*{\TheAuthor}{Manuel Schlund}
\newcommand*{\TheColloquiumDate}{2 August 2021}
\newcommand*{\TheDegree}{\foreignlanguage{ngerman}{%
    Doktor der Naturwissenschaften (\Dr{} rer.\@ nat.\@)}%
}
\newcommand*{\TheInstitute}{Institute of Environmental Physics (IUP)}
\newcommand*{\TheKeywords}{%
  Climate Modeling\sep Climate Model Weighting\sep Climate Projections\sep %
  Climate Sensitivity\sep Emergent Constraints\sep Machine Learning%
}
\newcommand*{\TheLanguage}{en-US}
\newcommand*{\TheLicense}{%
  Creative Commons Attribution 4.0 International License%
}
\newcommand*{\TheLicenseLink}{https://creativecommons.org/licenses/by/4.0/}
\newcommand*{\TheMonth}{August 2021}
\newcommand*{\ThePrimaryExaminer}{\Prof{} \Dr{} Veronika Eyring}
\newcommand*{\TheSecondaryExaminer}{\Prof{} \Dr{} Pierre Gentine}
\newcommand*{\TheSubmissionDate}{13 April 2021}
\newcommand*{\TheThesis}{Doctoral Dissertation}
\newcommand*{\TheTitle}{%
  Constraining Uncertainties in Multi-Model Projections of the Future Climate
  with Observations%
}
\newcommand*{\TheUniversity}{University of Bremen}